---
title: "Project_Group11"
output: html_notebook
---

#installing packages
install.packages("readxl")
install.packages("tidyr")

#attaching module libraries
library(dplyr)
library(readxl)
library(tidyr)

```{r}
#-------------------------[ DATA PREPROCESSING STEPS ]--------------------------
#--------------------------[ 1. IMPORTING DATASET ]-----------------------------

# Read data
df <- read_excel("/Users/tataurosetta/Downloads/rStudio Working Directory/RStudio Datasets/AirQualityUCI.xlsx")

#viewing default dataset
View(df)

# Checking datatypes of individual columns
str(df)

# Checking summary of the dataset
summary(df)

# Check for missing values
sum(is.na(df))

#--------------------------[ 2. HANDLING MISSING DATA ]-------------------------

# Changing -200 value to null value
df[df == -200] = NA

# Checking summary of the dataset after replacing -200 (or missing values)
#with NA 
summary(df)

#Viewing dataset after replacing -200 (or missing values) with NA
View(df)

#---[ 3. Replace the missing data with the average of the feature in which the data
#is missing: ]------------------------------------------------------------------

# Imputing mean for NA values
df <- df %>% 
  mutate_if(is.numeric, ~replace_na(.,mean(., na.rm = TRUE)))

# Checking summary of the dataset after replacing NA with mean of data
summary(df)

View(df)

```

# After the necessary steps taken in data preprocessing above,
# we are performing visualizations on the features of the AirQualityIndex data

# Below is the code for plotting a box plot on the features

```{r}

#boxplot

#defining colors for all the elements
myColors <- c('steelblue','red','orange','yellow','green','blue','gold',
              'violet','purple','pink')

#defining legend labels
elements <- c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", "NOx", "PT08.S3",
              "NO2", "PT08.S4", "PT08.S5")

#------------[ plotting the side-side box plot on a normal scale ]--------------

boxplot_normal_scale <- boxplot(df$"CO(GT)", df$"PT08.S1(CO)", 
            df$"NMHC(GT)", df$"C6H6(GT)", df$"PT08.S2(NMHC)", 
            df$"NOx(GT)", df$"PT08.S3(NOx)", df$"NO2(GT)",
            df$"PT08.S4(NO2)", df$"PT08.S5(O3)", names=elements, 
            horizontal = FALSE, 
            main = "Box Plot for all elements (Normal Scale)",
            xlab = "Elements", 
            ylab = "Hourly Measured Concentration in microg/m^3 & ppb",
            cex.axis = 0.59, cex.lab = 1.0, col=myColors, outcex = 0.62)

#defining the legend of the plot
legend("topleft", legend = c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", 
                  "NOx", "PT08.S3", "NO2", "PT08.S4", "PT08.S5"),
                  col = c('steelblue','red','orange','yellow','green','blue',
                  'gold','violet','purple','pink'), bty = "o", 
                  pch=20 , pt.cex = 3, cex = 0.7, horiz = FALSE, inset=0)



```
#  In the above diagram, a side-by-side box plot has been plotted for the features
#  from our data, more precisely only the elements sensed using metal oxide 
#  chemical sensory devices or sensors. The features that are relevant to this box
#  plot are CO(GT) [ Carbon Monoxide ], PT08.S1(CO) [ Tin Oxide (CO Nominally Targeted) ], 
#  NMHC(GT) [Nonmethane hydrocarbons ( NMHC, such as ethane, ethene, propane, propene, and isoprene ) ],
#  C6H6(GT) [ Benzene ], PT08.S2(NMHC) [ Titania ( NMHC Nominally Targeted)], 
#  NOx(GT) [ Nitric Oxide ], PT08.S3(NOx) [ Tungsten Oxide (NOx Nominally Targeted)],
#  NO2(GT) [ Nitrogen Dioxide ], PT08.S4(NO2) [ Tungsten Oxide (NO2 Nominally Targeted) ],
#  PT08.S5(O3) [ Indium Oxide (O3 Nominally Targeted) ]. The features are measured in
#  microg/m^3. As the box plot has been drawn, the first observations are telling us
#  there are a lot of outliers for each and every element/feature/column. Also elements
#  like CO and NO2 are visually undecipherable due to the scale on the y-axis. So before any
#  further interpretations are made, we need to remove the outliers and add log to the scale
#  for convenience in understanding the individual box distrbitutions for each element.

```{r}

#defining colors for all the elements
myColors <- c('steelblue','red','orange','yellow','green','blue','gold',
              'violet','purple','pink')

#defining legend labels
elements <- c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", "NOx", "PT08.S3",
              "NO2", "PT08.S4", "PT08.S5")

#---------[ plotting the side-side box plot on the logarithmic scale ]----------

boxplot_log_scale <- boxplot(df$"CO(GT)", df$"PT08.S1(CO)", 
            df$"NMHC(GT)", df$"C6H6(GT)", df$"PT08.S2(NMHC)", 
            df$"NOx(GT)", df$"PT08.S3(NOx)", df$"NO2(GT)",
            df$"PT08.S4(NO2)", df$"PT08.S5(O3)",
            log = "y", names=elements, horizontal = FALSE, 
            main = "Box Plot for all elements (Logarithmic scale)",
            xlab = "Elements", 
            ylab = "Hourly Measured Concentration in microg/m^3 & ppb",
            cex.axis = 0.59, cex.lab = 1.0, col=myColors, outcex = 0.62)

#defining the legend of the plot
legend("topleft", legend = c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", 
                  "NOx", "PT08.S3", "NO2", "PT08.S4", "PT08.S5"),
                  col = c('steelblue','red','orange','yellow','green','blue',
                  'gold','violet','purple','pink'), bty = "o", 
                  pch=20 , pt.cex = 3, cex = 0.7, horiz = FALSE, inset=0)

```
# Here, in this side-by-side box plot, we have added log to the y-scale for better
# understanding of individual box distribututions ( specially for elements like -
# CO(GT), NMHC(GT), C6H6(GT) and NO2(GT) ). As we can see, the box distributions are
# now visible for the above mentioned elements. However, the outliers are still present,
# and which we have eliminated from the plot using the code below and have displayed
# the final side-by-side box plot of our data for interpretations:

```{r}


#defining colors for all the elements
myColors <- c('steelblue','red','orange','yellow','green','blue','gold',
              'violet','purple','pink')

#defining legend labels
elements <- c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", "NOx", "PT08.S3",
              "NO2", "PT08.S4", "PT08.S5")

#---------[ plotting the side-side box plot without the outliers (only from)
#the plot ]---------------------------------------------------------------------
  
#Vertical 
  
boxplot_no_outliers <- boxplot(df$"CO(GT)", df$"PT08.S1(CO)", 
            df$"NMHC(GT)", df$"C6H6(GT)", df$"PT08.S2(NMHC)", 
            df$"NOx(GT)", df$"PT08.S3(NOx)", df$"NO2(GT)",
            df$"PT08.S4(NO2)", df$"PT08.S5(O3)",
            log = "y", names=elements, horizontal = FALSE, 
            main = "Box Plot for all elements (Logarithmic scale) (Outliers Removed)",
            xlab = "Elements", 
            ylab = "Hourly Measured Concentration in microg/m^3 & ppb",
            cex.axis = 0.59, cex.lab = 1.0, col=myColors,
            outcex = 0.62, outline = FALSE)

#defining the legend of the plot
legend("topleft", legend = c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", 
                  "NOx", "PT08.S3", "NO2", "PT08.S4", "PT08.S5"),
                  col = c('steelblue','red','orange','yellow','green','blue',
                  'gold','violet','purple','pink'), bty = "o", 
                  pch=20 , pt.cex = 3, cex = 0.7, horiz = FALSE, inset=0)


#---------[ plotting the side-side box plot without the outliers (only from)
#the plot ]---------------------------------------------------------------------
  
#Horizontal  
  
boxplot_no_outliers <- boxplot(df$"CO(GT)", df$"PT08.S1(CO)", 
            df$"NMHC(GT)", df$"C6H6(GT)", df$"PT08.S2(NMHC)", 
            df$"NOx(GT)", df$"PT08.S3(NOx)", df$"NO2(GT)",
            df$"PT08.S4(NO2)", df$"PT08.S5(O3)",
            log = "x", names=elements, horizontal = TRUE, 
            main = "Box Plot for all elements (Logarithmic scale) (Outliers Removed)",
            xlab = "Hourly Measured Concentration in microg/m^3 & ppb", 
            ylab = "Elements",
            cex.axis = 0.59, cex.lab = 1.0, col=myColors,
            outcex = 0.62, outline = FALSE)

#defining the legend of the plot
legend("topleft", legend = c("CO", "PT08.S1", "NMHC", "C6H6", "PT08.S2", 
                  "NOx", "PT08.S3", "NO2", "PT08.S4", "PT08.S5"),
                  col = c('steelblue','red','orange','yellow','green','blue',
                  'gold','violet','purple','pink'), bty = "o", 
                  pch=20 , pt.cex = 3, cex = 0.7, horiz = FALSE, inset=0)


```
#    In our final side-by-side box plot, all the outliers have been removed and the
#    following interpretations are made:

# For our final visualization, we have plotted a box plot on the relevant features, elements in our case, to give you an idea 
# about the concentrations measured hourly in micrograms in the air quality of the concerned location. 

# at first, the scale for which the box plot was to be plotted gave us distributions for some elements which could 
# not be deciphered due to scale range. Hence we took a logarthimic scale to properly identify the distributions of all elements
# and compare their values.

# I'll start off with some of the statistical interpretations of the given box plot. 

# 1. The first observation of the plot is that the dispersion for the elements [ CO(GT), C6H6(GT), and NOx(GT) ]
#    is considerably greater than the other elements. (The interquartile range, as shown by the box lengths, is longer ).

# 2. The second interpretation is that the Measure Concentration for the elements [ CO(GT), C6H6(GT), and NOx(GT) ]
#    are significantly spanned out a lot more than for the elements [ NMHC(GT), PT08.S1(CO), PT08.S2(NMHC), PT08.S3(NOx),
#    and PT08.S4(NO2) ] and slightly greater than for the elements [ NO2(GT) and PT08.S5(O3)] i.e., the median, 
#    the lower quartile and the upper quartile.

# 3. The longer boxes of [ CO(GT), C6H6(GT), and NOx(GT) ] represents a greater interquartile range indicating
#    the data not be a lot reliable. Since the interquartile range represents 50% of the data closest to the median,
#    a greater range in this section of the plot suggests that the median may not be a great indicator of central
#    tendency.

# 4. Also the longer whiskers of the elements [ CO(GT), C6H6(GT), and NOx(GT) ] means that the data for these 
#    features has a longer range in terms of the entire sample which is also a indicatior of less reliablity.

# 5. The long whiskers of the elements [ NO2(GT) and PT08.S5(O3) ] also means that the data for these features
#    may indicate low reliability but as the boxes of these elements are not long, we can say it's more reliable
#    than the data recorded for the elements [ CO(GT), C6H6(GT), and NOx(GT) ] as ultimately, longer whiskers are
#    lesser of a concern than longer boxes. A broad range of possibilities but a strong likelihood of central values
#    is more reliable to use for prediction than a moderate overall range with little concentration at the median.

# 6. The box for the elements [ CO(GT) and C6H6(GT) ] skewed. It is clearly left-skew, with a long tail 
#    to the left (high values), and also by the fact that the left part of the box (median to upper quartile)
#    is longer than the right part.

# 7. The box for the elements [ C6H6(GT) and PT08.S5(O3) ] is not symmetric, and its pattern of skewness is
#    slightly left-skew. It shows that the left whisker is longer than the right whisker and the left part of
#    the box is slightly longer than the right part.

# 8. The observation that can be inferred from the above plot is that the elements PT08.S1(CO) 
#    [ Tin Oxide (CO Nominally Targeted) ], PT08.S2(NMHC) [ Titania ( NMHC Nominally Targeted)], PT08.S3(NOx)
#    [ Tungsten Oxide (NOx Nominally Targeted)], NO2(GT) [ Nitrogen Dioxide ], PT08.S4(NO2) 
#    [ Tungsten Oxide (NO2 Nominally Targeted) ], PT08.S5(O3) [ Indium Oxide (O3 Nominally Targeted) ] are
#    found to be present in high concentrations at the location where the data was collected (or the elements
#    sensed using the metal oxide sensors). These mentioned elements are really hazardous to both living beings
#    and the climate and sadfety measures to reduce there concentrations needs to be taken immediately.

